# pre-class


Make sure you commit this often with meaningful messages. 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Read the source code for each of the following three functions, puzzle out what they do, and then brainstorm better names.

```{r}

# A better name would be compare_prefix
# The function returns FALSE if the prefix argument is not the prefix of the string argument and TRUE otherwise
compare_prefix <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix
}

# A better name would be remove_last
# The function takes away the last element of a vector
remove_last <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}

# A better name would be recycle_vec
# The function recycles the y vector for the length of the x vector
recycle_vec <- function(x, y) {
  rep(y, length.out = length(x))
}

```

2. Compare and contrast rnorm() and MASS::mvrnorm(). How could you make them more consistent?

*The functions rnorm() and mvrnorm() are similar in that they sample any specified amount of values from the normal distribution with a chosen mean and spread. The biggest difference is that rnorm() is limited to a single variable while mvrnorm() can consider multiple variables. Another difference comes in how the user inputs the spread. rnorm() takes the standard deviation, while mvrnorm() takes the covariance. I would make them consistent by making the spread in rnorm() be the variance rather than the standard deviation.*


3. Use `lapply()` and an anonymous function to find the coefficient of variation (the standard deviation divided by the mean) for all columns in the mtcars dataset. 

```{r}

# Used lapply to apply an anonymous function for all columns in mtcars
lapply(mtcars, function(x) sd(x)/mean(x))

```

4. Use vapply() to:
    a. Compute the standard deviation of every column in a numeric data frame.
    
```{r}

# Used vapply to apply sd to columns in the mtcars numeric data frame
vapply(mtcars, sd, numeric(1))

```

    b. Compute the standard deviation of every numeric column in a mixed data frame. (Hint: you’ll need to use vapply() twice.)

```{r}

# Used vapply to subset numeric columns and piped into the sd function
flights[vapply(flights, is.numeric, logical(1))] %>% vapply(sd, numeric(1))

```
