# pre-class


Make sure you commit this often with meaningfull messages. 


### Getting Started


We will work with the dataset called [gapminder](https://github.com/jennybc/gapminder), this is a cleaned up version from [Gapminder Data](http://www.gapminder.org/data/). Gapminder contains a lot of great data on all of the nations of the world. We first need to install the gapminder package in R. 

```
install.packages("gapminder")
```



```
library(dplyr)
library(gapminder)
gapminder
```



### Pre-Class Problems

Use **dplyr** functions to address the following questions:

1. How many unique countries are represented per continent?

```{r}
# Use dplyr functions to count number of distinct countries by continent
gapminder %>% group_by(continent) %>% distinct(country) %>% count()
```

2. Which European nation had the lowest GDP per capita in 1997? 

```{r}
# First filter to Eurpoe and 1997 then arrange by gdpPercap
gapminder %>% filter(continent == "Europe" & year == 1997) %>% arrange(gdpPercap)
```

*Albania was the Eurpoean nation with the lowest GDP per capita in 1997 with 3193.*

3. According to the data available, what was the average life expectancy across each continent in the 1980s?

```{r}
# First filter to 1980s then group by continent before finding mean
gapminder %>% filter(year < 1990 & year > 1979) %>% group_by(continent) %>%
  summarize(avg_life = mean(lifeExp))
```

4. What 5 countries have the highest total GDP over all years combined?

```{r}
# Create new variable with new GDP column
gapminder1 <- mutate(gapminder, GDP = gdpPercap * pop)

# Group by country, sum GDP through all years, then arrange in descending order
gapminder1 %>% group_by(country) %>% summarize(totalGDPAllYears = sum(GDP)) %>% arrange(desc(totalGDPAllYears))
```

*The 5 countries with the highest total GDP over all years combined are the US, Japan, China, Germany, and the UK.*

5. What countries and years had life expectancies of _at least_ 80 years? _N.b. only output the columns of interest: country, life expectancy and year (in that order)._

```{r}
# Filter to life expectancy of 80 or greater, then use select for desired columns and years
gapminder %>% filter(lifeExp >= 80) %>% select(country, lifeExp, year)
```

6. What 10 countries have the strongest correlation (in either direction) between life expectancy and per capita GDP?

```{r}
# Group by country and then use the cor function to find correlation between life expectancy and GDP per capita
gapminder %>% group_by(country) %>% summarize(correlation = cor(lifeExp, gdpPercap)) %>% arrange(desc(correlation))
```

7. Which combinations of continent (besides Asia) and year have the highest average population across all countries? _N.b. your output should include all results sorted by highest average population_. With what you already know, this one may stump you. See [this Q&A](http://stackoverflow.com/q/27207963/654296) for how to `ungroup` before `arrange`ing. This also [behaves differently in more recent versions of dplyr](https://github.com/hadley/dplyr/releases/tag/v0.5.0).
8. Which three countries have had the most consistent population estimates (i.e. lowest standard deviation) across the years of available data? 
9. Subset **gm** to only include observations from 1992 and store the results as **gm1992**. What kind of object is this?
10. Which observations indicate that the population of a country has *decreased* from the previous year **and** the life expectancy has *increased* from the previous year? See [the vignette on window functions](https://cran.r-project.org/web/packages/dplyr/vignettes/window-functions.html).
