# pre-class


Make sure you commit this often with meaningful messages. 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```








Standardizing a variable means subtracting the mean, and then dividing by the standard deviation. Let’s use a loop to standardize the numeric columns in the [Western Collaborative Group Study](https://clinicaltrials.gov/ct2/show/NCT00005174). This study began in 1960 with 3154 men ages 39-59, who were employed in one of 11 California based companies. They were followed until 1969 during this time, 257 of these men developed coronary heart disease (CHD). You can read this data in with the code below. You can access this dataset with the following code:

```{R}

suppressMessages(library(foreign))
wcgs <- read.dta("https://drive.google.com/uc?export=download&id=0B8CsRLdwqzbzYWxfN3ExQllBQkU")
```

The data has the following variables:



WCGS has the following variables:

-----------------------------------------------------------
Name    Description
------- -------------------------------------------
id      Subject identification number

age     Age in years

height  Height in inches

weight  Weight in lbs.

sbp     Systolic blood pressure in mm 

dbp     Diastolic blood pressure in mm Hg

chol    Fasting serum cholesterol in mm 

behpat  Behavior

  1       A1

  2       A2

  3       B3

  4       B4

ncigs   Cigarettes per day

dibpat  Behavior

1       type A

2       type B

chd69   Coronary heart disease

1       Yes

0       no

typechd Type of CHD

1       myocardial infarction or death

2       silent myocardial infarction

3       angina perctoris

time169 Time of CHD event or end of follow-up

arcus   Arcus senilis

0       absent

1       present

bmi     Body Mass Index
-----------------------------------------------------------
```{r}
# Call dplyr
#install.packages("dplyr")
library(dplyr)

```



### Question 1: Standardize Function

A. Create a function called standardize.me() that takes a numeric vector as an argument, and returns the standardized version of the vector. 


```{r}

# Create standardize function by subtracting mean and dividing by sd
standardize.me <- function(x) {
  result <- ((x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE))
  return(result)
}

# See that the function does as intended
x <- c(1:5)
standardize.me(x)
```

B. Assign all the numeric columns of the original WCGS dataset to a new dataset called WCGS.new.

```{r}


# Use select_if to extract numeric columns
WCGS.new <- select_if(wcgs, is.numeric)


head(WCGS.new)

```

C. Using a loop and your new function, standardize all the variables WCGS.new dataset.


```{r}

# Replace values with standardized values column-wise
for(i in 1:ncol(WCGS.new)){
  WCGS.new[,i] = standardize.me(WCGS.new[,i])
}

head(WCGS.new)
```


D. What should the mean and standard deviation of all your new standardized variables be? Test your prediction by running a loop

```{r}

# I would have predicted the means to be 0 and standard deviations to be 1

# Find mean and standard deviation column-wise
for(i in 1:ncol(WCGS.new)) {
  print(mean(WCGS.new[,i], na.rm = TRUE))
  print(sd(WCGS.new[,i], na.rm = TRUE))
}
```



### Question 2: Looping to Calculate

A. Using a loop, calculate the mean weight of the subjects separated by the type of CHD they have.

```{r}

# Create a vector of the unique types of CHD and put them into a data frame with initial means 0
chd_types <- as.vector(unlist(unique(wcgs["typchd69"])))
means <- c(0,0,0,0)
chd_means_df <- data.frame(chd_types, means)

# Use a for loop and subsetting to find the mean for each type of CHD and complete the data frame
for(i in 1:length(chd_types)) {
  chd_means_df[i,2] <- mean(wcgs[which(wcgs$typchd69 == chd_types[i]), "weight"], na.rm = TRUE)
}

# Show the results
chd_means_df

```

B. Now do the same thing, but now don’t use a loop

```{r}

# Much easier and more intuitive than part A to arrive at the same result
wcgs %>% group_by(typchd69) %>% summarize(mean_weight = mean(weight, na.rm = TRUE))

```


