# pre-class


Make sure you commit this often with meaningful messages. 




### Question 1:

Using a loop, print the integers from 1 to 50. 

```{r}

# Loops through 1 to 50 and prints each iteration
for(i in 1:50){
  print(i)
}

```

### Question 2:

A.  Using a loop, add all the integers between 0 and 1000.

```{r}

# Sums through 0 to 1000 and prints the final result
x <- 0
for(i in 0:1000){
  x <- x + i
}
print(x)

```

B. Now, add all the EVEN integers between 0 and 1000 (hint: use seq())

```{r}

# Sums the evens from 0 to 1000 and prints the final result
x <- 0
for(i in seq(0, 1000, by = 2)){
  x <- x + i
}
print(x)

```

C. Now, repeat A and B WITHOUT using a loop.

```{r}
# Part A
x <- c(0:1000)
sum(x)

#Part B
y <- seq(0, 1000, by = 2)
sum(y)

```

### Question 3:

Here is a dataframe of survey data containing 5 questions :

```{r, eval=FALSE}
survey <- data.frame(
                     "participant" = c(1, 2, 3, 4, 5, 6),
                     "q1" = c(5, 3, 2, 7, 11, 0),
                     "q2" = c(4, 2, 2, 5, -10, 99),
                     "q3" = c(-4, -3, 4, 2, 9, 10),
                     "q4" = c(-30, 5, 2, 23, 4, 2),
                     "q5" = c(88, 4, -20, 2, 4, 2)
                     )
```
The response to each question should be an integer between 1 and 5. Obviously, we have some bad values in the dataframe. The goal of this problem is to fix them.

A. Using a loop, create a new dataframe called survey.clean where all the invalid values (those that are not integers between 1 and 5) are set to NA.

```{r}

# Uses a nested for loop to scan through each position in survey, 
# replacing any values that are not valid
for(i in 1:nrow(survey)){
  for(j in 1:ncol(survey)){
    if(survey[i,j] > 5 | survey[i,j] < 1){
      survey[i,j] <- NA
    }
  }
}

print(survey)
```

B. Now, again using a loop, add a new column to the dataframe called “invalid.answers” that indicates, for each participant, how many bad answers they gave.

```{r}

# First create a new column called "invalid.answers" with all zeros
survey$invalid.answers <- 0

survey

# Use a nested for loop that adds 1 to "invalid.answers" each time an NA
# is encountered
for(i in 1:nrow(survey)){
  for(j in 1:ncol(survey)){
    if(is.na(survey[i,j])){
      survey[i, 7] = survey[i, 7] + 1
    }
  }
}

survey

```

